---
title: "rmaxima - an interface to maxima for R"
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

Development version of the yet-to-be-released rmaxima CRAN package. rmaxima provides an interface to the powerful and fairly complete maxima computer algebra system

This repository is a fork from the original version created by Kseniia Shumelchyk and Hans W. Borcher [shumelchyk/rmaxima](https://github.com/shumelchyk/rmaxima), which is currently not maintained.

# Installation

## Requirements
- this package currently only works under Linux (and has not been tested under MacOS)
- you can install it without having Maxima installed, but obviously need to install Maxima in order to use it.

## Steps

If you want to install the latest version (currently the only one available) install the R package `drat` first and add this github account as a repo:

```{r, eval = FALSE}
install.packages("drat")
drat::addRepo("rcst")
```

Now you can easily install it the usual way:

```{r eval = FALSE}
install.packages("rmaxima")
```
