---
title: "rim - R's interface to Maxima"
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

[![](https://cranlogs.r-pkg.org/badges/rim)](https://cran.rstudio.com/web/packages/rim/index.html)

[![](https://cranlogs.r-pkg.org/badges/grand-total/rim)](https://cran.rstudio.com/web/packages/rim/index.html)

Development version of CRAN package. `rim` provides an interface to the powerful and fairly complete maxima computer algebra system

This repository is a fork from RMaxima, the original version created by Kseniia Shumelchyk and Hans W. Borcher [shumelchyk/rmaxima](https://github.com/shumelchyk/rmaxima), which is currently not maintained.

# Installation

```{r, eval = FALSE}
install.packages("rim")
```


## Requirements
- Maxima (https://sourceforge.net/projects/maxima/, tested with versions >=5.42.1), needs to be on PATH


## Latest Version

If you want to install the latest version install the R package `drat` first and add this github account as a repo:

```{r, eval = FALSE}
install.packages("drat")
drat::addRepo("rcst")
```

Now you can easily install it the usual way:

```{r eval = FALSE}
install.packages("rim")
```

# Usage

This section only demonstrate using the packages R-function directly accessible to the user. On how to use the package's `knitr` engine see [this page](https://rcst.github.io/rim/).

```{r, eval = FALSE}
library(rim)
```

```{r}
maxima.start(restart = TRUE)
maxima.get("1+1;")
r <- maxima.get("sum(1/x^2, x, 1, 10000)")
print(r)
iprint(r)
maxima.isInstalled()
maxima.version()
maxima.getformat()
maxima.setformat("text2d")
maxima.apropos("int")
maxima.get("integrate(1 / (x^4 + 1), x);")
maxima.get("jacobian( [alpha / (alpha + beta), 1 / sqrt(alpha + beta)], [alpha, beta] )")
l <- maxima.get("%;")
unclass(l)
maxima.load("abs_integrate")
maxima.stop()
```
