---
title: "rim - R's interface to Maxima"
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

Development version of the yet-to-be-released CRAN package. rim provides an interface to the powerful and fairly complete maxima computer algebra system

This repository is a fork from RMaxima, the original version created by Kseniia Shumelchyk and Hans W. Borcher [shumelchyk/rmaxima](https://github.com/shumelchyk/rmaxima), which is currently not maintained.

# Installation

## Requirements
- Maxima

## Steps

If you want to install the latest version (currently the only one available) install the R package `drat` first and add this github account as a repo:

```{r, eval = FALSE}
install.packages("drat")
drat::addRepo("rcst")
```

Now you can easily install it the usual way:

```{r eval = FALSE}
install.packages("rim")
```

# Usage

This section only demonstrate using the packages R-function directly accessible to the user. On how to use the package's `knitr` engine see [this page](https://rcst.github.io/rim).

```{r, eval = FALSE}
library(rim)
```

```{r}
maxima.start(restart = TRUE)
maxima.get("1+1;")
maxima.get("y: 0.5 * t^2 * g + g*t;")
maxima.get("diff(y, t);")
maxima.stop()
```
