---
title: "rim - R's interface to Maxima"
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

[![](https://cranlogs.r-pkg.org/badges/rim)](https://CRAN.R-project.org/package=rim) 
[![](https://cranlogs.r-pkg.org/badges/grand-total/rim)](https://CRAN.R-project.org/package=rim)

Development version of CRAN package. `rim` provides an interface to the powerful and fairly complete maxima computer algebra system

This repository is a fork from RMaxima, the original version created by Kseniia Shumelchyk and Hans W. Borcher [shumelchyk/rmaxima](https://github.com/shumelchyk/rmaxima), which is currently not maintained.

I have forked this project because I was not able to make the original one work, my objectives are brief but perhaps useful for future navigators, they are as follows:
*Fix bugs that didn't let me use this project in conditions.
*To create new functions to be able to use R together with Maxima in a simpler way (to be able to pass values from R to Maxima in a comfortable way, for example).
*Eliminate all the instances of S6 and modify them by S3, since I think that it is not an OOP system that really contributes of great measure, complicating the debugability of the project.

In short, my goal is to make a project that works for me, making public my modifications in case they are useful to others.

Translated with DeepL.com (free version)







# Installation

```{r, eval = FALSE}
install.packages("rim")
```


## Requirements
- Maxima (https://sourceforge.net/projects/maxima/, tested with versions >=5.42.1), needs to be on PATH

## Latest Version

If you want to install the latest version install the R package `remotes` first and then install the package from this github repo:

```{r, eval = FALSE}
install.packages("remotes")
remotes::install_github("ManuelThings/rim")
```

# Usage

This section only demonstrates the package's R-function directly accessible to the user. On how to use the package's `knitr` engine see [this page](https://rcst.github.io/rim/).

```{r, eval = FALSE}
library(rim)
```

```{r}
devtools::load_all(".")
maxima.start(restart = TRUE)
maxima.get("1+1;")
r <- maxima.get("sum(1/x^2, x, 1, 10000)")
print(r)
iprint(r)
maxima.isInstalled()
maxima.version()
maxima.options
maxima.options(format = "ascii")
maxima.apropos("int")
maxima.get("integrate(1 / (x^4 + 1), x);")
maxima.get("jacobian( [alpha / (alpha + beta), 1 / sqrt(alpha + beta)], [alpha, beta] )")
l <- maxima.get("%;")
maxima.eval(l, code = TRUE, envir = list(alpha = 0.3, beta = 2.5))
unclass(l)
maxima.load("abs_integrate")
maxima.stop()
```
